Case for Support

### 2021-11-16 application/bonsai.tex
Updates from Overleaf

Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
programming language that emphasizes performance, flexibility, and ease-of-use,
allowing scientists with no previous programming experience to quickly develop
their own high-performance data acquisition and experimental control systems.
Bonsai combines a high-level event algebra for data streams with an integrated
development environment (IDE) and a strong library of plugins supporting
multiple hardware and software packages used by the biomedical research
community (Figures~\ref{fig:bonsai}a and~\ref{fig:bonsai}b).
%
A Bonsai graphical program consists of one or more source data streams
and several interconnected operators that transform input to output
datastreams (Figure~\ref{fig:bonsai}c).

Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have been
transformative to the progress of increasingly “data-rich” sciences. However,
the equivalent standardized software tools for data acquisition and control of
animal behaviour experiments are still lacking. Life science experiments demand
a combination of multiple instrumentation and control technologies, for both
behavioral and physiological investigations. The growing complexity on both the
amount of data that is collected, and the rich conditions under which behaviour
must be explored, place an increasing burden on experimenters to integrate
highly specialized equipment in unique configurations, while often lacking
expertise in the relevant engineering fields.

Bonsai has a large user base in the systems neuroscience community
(Figures~\ref{fig:bonsai}d-\ref{fig:bonsai}f).  In the last year alone, we estimate more than 1,000 new users have
started to incorporate Bonsai into their experimental protocols across the
world. The surprising rate of adoption of Bonsai in non-programmer experimental
labs highlights the need for accessible programming tools that enable
state-of-the-art technology but also allow researchers to stay in control and
change their experimental paradigms.  Many open-source software tools are
either inaccessible to non-programmers, or too constrained to be of general use
outside their narrow domain of application. Bonsai has been successful because
it has straddled this gap to some extent.

>>>>

Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
programming language that emphasizes performance, flexibility, and ease-of-use,
allowing scientists with no previous programming experience to quickly develop
their own high-performance data acquisition and experimental control systems.
Bonsai combines a high-level event algebra for data streams with an integrated
development environment (IDE) and a strong library of plugins supporting
multiple hardware and software packages used by the biomedical research
community (Figures~\ref{fig:bonsai}a and~\ref{fig:bonsai}b).
%
A Bonsai graphical program consists of one or more source data streams (e.g. neural activity, video, audio, sensors, etc)
and several interconnected operators that transform input to output
datastreams (Figure~\ref{fig:bonsai}c).

Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have been
transformative to the progress of increasingly “data-rich” sciences. However,
the equivalent standardized software tools for data acquisition and control of
animal behaviour neuroscience experiments are still lacking. Life science experiments demand
a combination of multiple instrumentation and control technologies, for both
behavioural and physiological investigations. The growing complexity on both the
amount of data that is collected, and the rich conditions under which behaviour
must be explored, place an increasing burden on experimenters to integrate
highly specialised equipment in unique configurations, while often lacking
expertise in the relevant engineering fields.

Bonsai has been in adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figures~\ref{fig:bonsai}d-\ref{fig:bonsai}f).  In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols. The surprising rate of adoption of Bonsai in non-programmer experimental
labs highlights the need for accessible programming tools that enable
state-of-the-art technology but also allow researchers to stay in control and flexibly
change their experimental paradigms.  Many open-source software tools are
either inaccessible to non-programmers, or too constrained to be of general use
outside their narrow domain of application. Bonsai has been successful because
it has straddled this gap.
